---
title: "hw4-solution-key"
author: "Nathaniel Grimes"
date: "2023-05-26"
output: 
  html_document:
    code_folding: show
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, 
                      echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(readr)
library(cowplot)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(rootSolve)
library(janitor)
library(stargazer)


#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation


```

# Estimated damage function

The following code estimates our quadratic damage function while omitting the intercept term 


```{r , include=TRUE}

## Load damage data and create the squared warming variable
damages_data <- read_csv("damages.csv") %>%
  clean_names()%>%
  mutate(warming2 = warming^2)

## Load climate trajectories
warming_data <- read_csv("warming.csv")%>%
  clean_names()

## Question 1
model_1 <- lm(damages ~ warming + warming2 -1, data = damages_data)

## Add fitted data 
damages_fitted = model_1$model %>%
  mutate(fitted=model_1$fitted.values) 

## Plot raw and fitted data
plot_1 <- ggplot(data=damages_fitted) + 
  geom_point(aes(x=warming,y=damages/1E12),color="red",alpha=.5) +
  geom_line(aes(x=warming,y=fitted/1E12),size=2)+
  theme_cowplot(14)+
  labs(x = "Warming (degrees C)", y = "Damages (trillions of dollars)")

```

<center>**Figure 1: Estimated quadratic damage function**</center>
```{r , fig.width = 5, fig.height = 4, eval=TRUE, echo=FALSE}
plot_1
```

# Predicting damages of warming scenarios
```{r , include=TRUE}

## Saving quadratic coefficients 
b_1 = model_1$coefficients[[1]]
b_2 = model_1$coefficients[[2]]

## Define pulse
pulse <- 35E9

## Predict damages 
warming_damages <- warming_data %>%
  mutate(damage_nopulse = b_1*warming_baseline + b_2*warming_baseline^2) %>%
  mutate(damage_pulse = b_1*warming_pulse + b_2*warming_pulse^2) %>%
  mutate(damage_diff = damage_pulse - damage_nopulse) %>%
  mutate(damage_diff_perton = damage_diff/pulse)

P2a <- ggplot(data=warming_damages) +
  geom_line(aes(x=year,y=damage_nopulse/1E12))+
  theme_cowplot(14)+
  labs(x = "Year", y = "Trillions of $", title = "A. Damages, no pulse")
P2b <- ggplot(data=warming_damages) +
  geom_line(aes(x=year,y=damage_pulse/1E12))+
  theme_cowplot(14)+
  labs(x = "Year", y = "Trillions of $", title = "B. Damages,  pulse")
P2c <- ggplot(data=warming_damages) +
  geom_line(aes(x=year,y=damage_diff/1E9))+
  theme_cowplot(14)+
  labs(x = "Year", y = "Billions of $", title = "C. Damage differences")
P2d <- ggplot(data=warming_damages) +
  geom_line(aes(x=year,y=damage_diff_perton))+
  theme_cowplot(14)+
  labs(x = "Year", y = "$ per ton", title = "D. Differences per ton")

grid.arrange(P2a, P2b, P2c, P2d, ncol=2, top ="Figure 2: Predicted damages")

```

# The social cost of carbon by discount rates

```{r , include=TRUE}

## Generate a range of discount rates 
# I look at discount rates from 0 to 5%

r <-  seq(0.01,.08,0.01)

ssc<-purrr::map(r,\(x) warming_damages %>% 
                     mutate(t=year-2021) %>% 
                     mutate(PVdiff_perton=damage_diff_perton/(1+x)^t) %>% #Discount
                      summarise(ssc=sum(PVdiff_perton),r=x) ) %>% 
             purrr::list_rbind()


## PV of damage differences per ton for different discount rates 
damages_PV <- lapply(1:length(r), function(x) 
  bind_cols(r= r[x], 
            PVdiff_perton = warming_damages$damage_diff_perton/(1+r[x])^(warming_damages$year-2021),
            year = warming_damages$year))%>%
  bind_rows()

## Summing the PVs per discount rate to get SCC

SCC <- damages_PV %>%
  group_by(r)%>%
  summarise(SCC = sum(PVdiff_perton))

## Ramsey rule discount rate

r_ramsey <- 0.001 + 2*0.01

plot_3 <- ggplot(data=SCC) +
  geom_line(aes(x=r,y=SCC),size=2)+
  theme_cowplot(14)+
  labs(x = "Discount rate", y = "Social cost of carbon ($ per ton)")+ 
  geom_vline(xintercept = r_ramsey, color="red") 

```

<center>**Figure 3: Social cost of carbon by discount rates **</center>
```{r , fig.width = 5, fig.height = 4, eval=TRUE, echo=FALSE}
plot_3
```


# The SCC under Ramsey Rule

The red line in Figure 3 shows the SCC of around $\$71$ at the Ramsey rule discount rate of 2.1\%.



# Decision making under uncertainty

```{r, include=TRUE}

# PV of damages under different policies and scenarios
policy_damages <- damages_PV %>%
  filter(r==0.02) %>%
  left_join(warming_data %>% select(year, warm_A1=warming_baseline), by= c("year"="year"))%>%
  mutate(warm_A2=1.5*warm_A1) %>%
  mutate(warm_Ba=warm_A1) %>%
  mutate(warm_Bb=1.29) %>%
  mutate(warm_B=pmin(warm_Ba,warm_Bb)) %>%
  mutate(damage_A1=(b_1*warm_A1 + b_2*warm_A1^2)/((1+0.02)^(year-2021))) %>%
  mutate(damage_A2=(b_1*warm_A2 + b_2*warm_A2^2)/((1+0.02)^(year-2021))) %>%
  mutate(damage_B=(b_1*warm_B + b_2*warm_B^2)/((1+0.02)^(year-2021))) 

# Sum of PV of damages between both policies
expected_value_A = .5*sum(policy_damages$damage_A1) + .5*sum(policy_damages$damage_A2)
value_B = sum(policy_damages$damage_B)

```

The expected present value of damages under Policy A are of \$`r format(round(expected_value_A/1E12), big.mark=",")` trillions. The expected present value of damages under Policy B are of \$`r format(round(value_B/1E12), big.mark=",")` trillions.

When society is risk neutral, we know society's utility function $U(m)$ (I use $m$ to denote money as the parameter in the utility function) is linear, and that $E(U(m)) = U(E(m))$. In the case of the random outcome of Policy A, $A$, $E(U(A)) = U(E(A)) = U(-\$3,023)$. For the certain outcome of Policy B, $B$, $U(B) = U(-\$787 - X)$, were $X$ is the policy cost. 

```{r}
# Let's look at the cost of the policy from 0 to 5000 trillion
x= seq(0, 5000, length.out = 100)

cost_scenario = tibble(X=x, cost_b = 1620+x, cost_a=6314)

plot_4 <- ggplot(cost_scenario, aes(X, y=cost_b)) +
  geom_line() +
  geom_hline(yintercept = 6314, color = 'mistyrose4')  +
  theme_cowplot() +
  geom_text(data = tibble(X=c(3500, 3500), cost_b = c(3200, 5200), 
                          policy = c('Policy B','Policy A')), aes(label = policy)) +
  labs(y = 'Expected PV damage $ (trillions)', x= 'X ($)')
```



Suppose the cost were zero or $X=0$, then we know that a risk neutral society would prefer Policy B since $E(U(A)) = U(E(A)) = U(-\$3,023)<U(B) = U(-\$787)$. We know this since $U(m)$ is increasing in $w$, utility is always increasing in \$.



<center>**Figure 4: Total expected cost of policy A and B under a range of policy B cost, X. **</center>
```{r , fig.width = 5, fig.height = 4, eval=TRUE, echo=FALSE}
plot_4
```



Society will prefer Policy B until $-\$3,023 = -\$787 - X$ or $X = \$2,236$.


# Risk averse society

If society is risk averse, then $E(U(m)) < U(E(m))$. This implies that when $X=\$2,236$, E(U(A))< U(E(A)) = U(B). Therefore, since E(U(A))<U(B) when $X=\$2301$, we can say that when society is risk averse, they would still prefer Policy B to A for costs greater than \$2,236. However, we cannot put an upper bound on those costs, since we would need to know more about the functional form of $U(m)$. 
